---
title: 'MA: Bias when adding trial'
author: "Martin Law"
date: '2022-06-06'
output: rmarkdown::html_vignette
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  dev='png',
  fig.path="figures/")
```

```{r setup}
library(pwb)
library(ggplot2)
```

### Trial design with no precision weighted bias

We identify a trial design with substantial unadjusted bias and no precision weighted bias. We obtain bias for a meta-analysis before and after including this design in the meta-analysis. We expect no increase in (absolute) bias.


Find a Simon design with p~0~=0.5, p~1~=0.6, type-I error-rate 0.05, power
0.9:

```{r find_design}
all.designs <- clinfun::ph2simon(pu=0.5,
                                 pa=0.6,
                                 ep1=0.05,
                                 ep2=0.1,
                                 nmax=500)
# Choose optimal Simon design:
opt.index <- which.min(all.designs$out[, "EN(p0)"])
simon.des <- all.designs$out[opt.index, ]
```


This design has some unadjusted bias and no precision weighted bias when true response probability theta=0.5:

```{r, find bias single }
  nsims <- 1e5
  theta <- 0.5
  set.seed(16)
  single.simon <- pwbSimon(theta=theta,
                                des=simon.des,
                                nsims=nsims)
  round(single.simon$ests[3:4, 2:4], 5)
```

There is unadjusted bias and no precision weighted bias.


Now simulate all other meta-analysis studies and find summary estimate for each replication. These studies all have sample size equal to the maximum sample size of the Simon design.

```{r}
ma.results <- simMeta(N=simon.des["n"], theta=0.5, n.studies=4, nsims=nsims)
wtdmeans.ex <- rep(NA, nsims)
for(i in 1:nsims) wtdmeans.ex[i] <- weighted.mean(x=ma.results$theta.hat[i, ], w=1/ma.results$se[i, ]^2)

```
Now add the replications of the Simon design to the meta-analysis and find summary estimates again:

```{r}
combined.results <- ma.results
combined.results$theta.hat <- cbind(single.simon$results$theta.hat, ma.results$theta.hat)
combined.results$se <- cbind(single.simon$results$se, ma.results$se)

wtdmeans.inc <- rep(NA, nsims)
for(i in 1:nsims) wtdmeans.inc[i] <- weighted.mean(x=combined.results$theta.hat[i, ], w=1/combined.results$se[i, ]^2)
```

Now find the bias in the meta-analysis, excluding and including the Simon design:

```{r}
bias.ex <- mean(wtdmeans.ex-theta)
bias.inc <- mean(wtdmeans.inc-theta)
bias.ex
bias.inc
abs(bias.inc)-abs(bias.ex)
```
Bias increases by almost an order of magnitude but still remains very small.



### Trial design with some precision weighted bias

We now do the converse: identify a trial with precision weighted bias but no unadjusted bias.

We start by finding a Simon design with p~0~=0.8, p~1~=0.9, type-I error-rate 0.05, power
0.9:

```{r}
all.designs2 <- clinfun::ph2simon(pu=0.8,
                                 pa=0.9,
                                 ep1=0.05,
                                 ep2=0.1,
                                 nmax=150)
# Choose optimal Simon design:
opt.index <- which.min(all.designs2$out[, "EN(p0)"])
simon.des2 <- all.designs2$out[opt.index, ]
```


Find bias for true response probability theta=0.65:

```{r, find bias single 2 }
# Find bias for this design, for a single true response probability theta:
  theta <- 0.65
  set.seed(16)
  single.simon2 <- pwbSimon(theta=theta,
                                des=simon.des2,
                                nsims=nsims)
  round(single.simon2$ests[3:4, 2:4], 5)
```
There is some unadjusted bias but the precision weighted bias is more than one order of magnitude greater.


Again, simulate all other meta-analysis studies and find summary estimate for each replication. These studies all have sample size equal to the maximum sample size of the Simon design directly above.

```{r}
ma.results2 <- simMeta(N=simon.des2["n"], theta=0.65, n.studies=4, nsims=nsims)
wtdmeans.ex2 <- rep(NA, nsims)
for(i in 1:nsims)    wtdmeans.ex2[i] <- weighted.mean(x=ma.results2$theta.hat[i, ], w=1/ma.results2$se[i, ]^2)
```

Now add replications of this second Simon design to a meta-analysis:

```{r}
combined.results2 <- ma.results2
combined.results2$theta.hat <- cbind(single.simon2$results$theta.hat, ma.results2$theta.hat)
combined.results2$se <- cbind(single.simon2$results$se, ma.results2$se)

# Summary estimate and bias (including Simon design):
wtdmeans.inc2 <- rep(NA, nsims)
for(i in 1:nsims) wtdmeans.inc2[i] <- weighted.mean(x=combined.results2$theta.hat[i, ], w=1/combined.results2$se[i, ]^2)

bias.ex2 <- mean(wtdmeans.ex2-theta)
bias.inc2 <- mean(wtdmeans.inc2-theta)
bias.ex2
bias.inc2
abs(bias.inc2)-abs(bias.ex2)
```
Bias increases only slightly, but we expect a larger increase.
