---
title: 'MA: Bias when adding trial'
author: "Martin Law"
date: '2022-06-06'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MA_add_trial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  dev='png',
  fig.path="figures/")
options(knitr.kable.NA = '---')
```

```{r setup}
librarian::shelf(martinlaw/pwb, ggplot2, kableExtra, flextable)
```

### Trial design with no precision weighted bias

We identify a trial design with substantial unadjusted bias and no precision weighted bias. We obtain bias for a meta-analysis before and after including this design in the meta-analysis. We expect no increase in (absolute) bias.


Find a Simon design with p~0~=0.5, p~1~=0.6, type-I error-rate 0.05, power
0.9:

```{r find_design}
all.designs <- clinfun::ph2simon(pu=0.5,
                                 pa=0.6,
                                 ep1=0.05,
                                 ep2=0.1,
                                 nmax=500)
# Choose optimal Simon design:
opt.index <- which.min(all.designs$out[, "EN(p0)"])
simon.des <- all.designs$out[opt.index, ]
```


This design has some unadjusted bias and no precision weighted bias when true response probability theta=0.5:

```{r find bias single }
  nsims <- 1e5
  theta <- 0.5
  set.seed(16)
  single.simon <- pwbSimon(theta=theta,
                                des=simon.des,
                                nsims=nsims)
  round(single.simon$ests[3:4, 2:4], 5)
  mcse.simon <- MCSEbias(single.simon$results$theta.hat)
```

There is unadjusted bias and no precision weighted bias.


Now simulate all other meta-analysis studies and find summary estimate for each replication. These studies all have sample size equal to the maximum sample size of the Simon design.

```{r MA}
ma.results <- simMeta(N=simon.des["n"], theta=0.5, n.studies=4, nsims=nsims)
wtdmeans.ex <- rep(NA, nsims)
for(i in 1:nsims) wtdmeans.ex[i] <- weighted.mean(x=ma.results$theta.hat[i, ], w=1/ma.results$se[i, ]^2)

```
Now add the replications of the Simon design to the meta-analysis and find summary estimates again:

```{r}
combined.results <- ma.results
combined.results$theta.hat <- cbind(single.simon$results$theta.hat, ma.results$theta.hat)
combined.results$se <- cbind(single.simon$results$se, ma.results$se)

wtdmeans.inc <- rep(NA, nsims)
for(i in 1:nsims) wtdmeans.inc[i] <- weighted.mean(x=combined.results$theta.hat[i, ], w=1/combined.results$se[i, ]^2)
```

Now find the bias in the meta-analysis, excluding and including the Simon design:

```{r}
bias.ex <- mean(wtdmeans.ex-theta)
bias.ex
unwtd.bias.ex <- mean(rowMeans(ma.results$theta.hat)) - theta # Bias using simple mean of estimates
unwtd.bias.ex

bias.inc <- mean(wtdmeans.inc-theta)
bias.inc
unwtd.bias.inc <- mean(rowMeans(combined.results$theta.hat)) - theta # Bias using simple mean of estimates
unwtd.bias.inc

abs(bias.inc)-abs(bias.ex)

# MCSE:
mcse.ex <- MCSEbias(wtdmeans.ex)
mcse.ex * 1e5
mcse.inc <- MCSEbias(wtdmeans.inc)
mcse.inc * 1e5

```

Bias increases by almost an order of magnitude but still remains very small.

```{r bias MCSE non-ADs}
# Bias of each non-AD study:
bias.nonAD <- colMeans(ma.results$theta.hat)-theta
# PWB
pwb.nonAD <- rep(NA, ncol(ma.results$theta.hat))
for(i in 1:ncol(ma.results$theta.hat)){
  pwb.nonAD[i] <- weighted.mean(x=ma.results$theta.hat[, i], w=1/ma.results$se[,i]^2) - theta
}
# MCSE of each non-AD study:
mcse.nonAD <- apply(ma.results$theta.hat, 2, MCSEbias)
```




```{r precision}
emp.SE <- apply(combined.results$theta.hat, 2, empSE) # emp SE of each column
precision <- 1/(emp.SE^2)
# Precision of Simon design as fraction of precision of all designs:
simon.frac <- precision[1]/sum(precision)

prec.simon <- precision[1]
ave.prec.non.ad <- mean(precision[-1])
# Precision of Simon:
prec.simon
# Mean precision of non-adaptive design studies:
ave.prec.non.ad
# Fraction of Simon vs mean of non-adaptive:
prec.frac <- prec.simon/ave.prec.non.ad
prec.frac

```
For simulation study 1, average precision of Simon as a fraction of the non-adaptive studies is `r round(prec.frac, 2)`.

#### What is the bias of each individual trial? 22nd July 2024




### Trial design with some precision weighted bias

We now do the converse: identify a trial with precision weighted bias but no unadjusted bias.

This is the case for the Simon design above when the true response probability is theta=0.3:

```{r, find bias single 2 }
# Find bias for this design, for a single true response probability theta:
  theta <- 0.3
  set.seed(16)
  single.simon2 <- pwbSimon(theta=0.3,
                                des=simon.des,
                                nsims=nsims)
  round(single.simon2$ests[3:4, 2:4], 5)
  mcse.simon2 <- MCSEbias(single.simon2$results$theta.hat)
```
There is some unadjusted bias but the precision weighted bias is more than one order of magnitude greater.


Again, simulate all other meta-analysis studies and find summary estimate for each replication. These studies all have sample size equal to the maximum sample size of the Simon design directly above.

```{r}
# Summary estimates (excluding Simon design):
ma.results2 <- simMeta(N=simon.des["n"], theta=theta, n.studies=4, nsims=nsims)
wtdmeans.ex2 <- rep(NA, nsims)
for(i in 1:nsims)    wtdmeans.ex2[i] <- weighted.mean(x=ma.results2$theta.hat[i, ], w=1/ma.results2$se[i, ]^2)

```

Now add replications of this second Simon design to a meta-analysis:

```{r}
combined.results2 <- ma.results2
combined.results2$theta.hat <- cbind(single.simon2$results$theta.hat, ma.results2$theta.hat)
combined.results2$se <- cbind(single.simon2$results$se, ma.results2$se)

# Summary estimates (including Simon design):
wtdmeans.inc2 <- rep(NA, nsims)
for(i in 1:nsims) wtdmeans.inc2[i] <- weighted.mean(x=combined.results2$theta.hat[i, ], w=1/combined.results2$se[i, ]^2)
```

Obtain bias excluding/including Simon:
```{r}
# Bias excluding:
bias.ex2 <- mean(wtdmeans.ex2-theta)
bias.ex2
unwtd.bias.ex2 <- mean(rowMeans(ma.results2$theta.hat)) - theta # Bias using simple mean of estimates
unwtd.bias.ex2


# Bias including:
bias.inc2 <- mean(wtdmeans.inc2-theta)
bias.inc2
unwtd.bias.inc2 <- mean(rowMeans(combined.results2$theta.hat)) - theta # Bias using simple mean of estimates
unwtd.bias.inc2

# Absolute bias, difference:
abs(bias.inc2)-abs(bias.ex2)

mcse.ex2 <- MCSEbias(wtdmeans.ex2)
mcse.ex2 * 1e5
mcse.inc2 <- MCSEbias(wtdmeans.inc2)
mcse.inc2 * 1e5
```
Absolute bias increases only slightly, but we expect a larger increase.


```{r bias MCSE non-ADs theta equals 0.3}
# Bias of each non-AD study:
bias.nonAD2 <- colMeans(ma.results2$theta.hat)-theta
# PWB
pwb.nonAD2 <- rep(NA, ncol(ma.results2$theta.hat))
for(i in 1:ncol(ma.results2$theta.hat)){
  pwb.nonAD2[i] <- weighted.mean(x=ma.results2$theta.hat[, i], w=1/ma.results2$se[,i]^2) - theta
}
# MCSE of each non-AD study:
mcse.nonAD2 <- apply(ma.results2$theta.hat, 2, MCSEbias)
```

Find the average precision of the non-adaptive design studies and the Simon design:

```{r precision 2}
emp.SE2 <- apply(combined.results2$theta.hat, 2, empSE) # emp SE of each column
precision2 <- 1/(emp.SE2^2)
# Precision of Simon design as fraction of precision of all designs:
simon.frac2 <- precision2[1]/sum(precision2)

prec.simon2 <- precision2[1]
ave.prec.non.ad2 <- mean(precision2[-1])
# Precision of Simon:
prec.simon2
# Mean precision of non-adaptive design studies:
ave.prec.non.ad2
# Fraction of Simon vs mean of non-adaptive:
prec.frac2 <- prec.simon2/ave.prec.non.ad2
prec.frac2
```

For simulation study 2, average precision of Simon as a fraction of the non-adaptive studies is `r round(prec.frac2, 2)`.

### Summary of biases and MCSEs


```{r summary, results='asis'}
simon.unwtd.bias <- c(single.simon$ests["All", "bias"], single.simon2$ests["All", "bias"])
simon.pw.bias <- c(single.simon$ests["All (precision-weighted)", "bias"], single.simon2$ests["All (precision-weighted)", "bias"])
ma.bias.ex <- c(bias.ex, bias.ex2)
ma.bias.inc <- c(bias.inc, bias.inc2)
ma.unwtd.bias.ex <- c(unwtd.bias.ex, unwtd.bias.ex2)
ma.unwtd.bias.inc <- c(unwtd.bias.inc, unwtd.bias.inc2) 
ma.mcse.ex <- c(mcse.ex, mcse.ex2)
ma.mcse.inc <- c(mcse.inc, mcse.inc2)
simon.mcse <- c(mcse.simon, mcse.simon2)
#abs.change.bias <- abs(abs(ma.bias.ex) - abs(ma.bias.inc)) NOT CORRECT
#fraction <- 100*abs.change.bias/abs(simon.unwtd.bias) # HENCE NOT CORRECT
#fraction.pwb <- 100*abs.change.bias/abs(simon.pw.bias) # HENCE NOT CORRECT
# ETC....
# 
# results <- data.frame(simon.unwtd.bias,
#                       simon.pw.bias,
#                       simon.mcse,
#                       abs.change.bias,
#                       fraction,
#                       fraction.pwb,
#                       ma.mcse.ex,
#                       ma.mcse.inc)
# results <- results*1e5
# results$fraction <- results$fraction/1e5
# results$fraction.pwb <- results$fraction.pwb/1e5
# results <- cbind(theta=c("0.5", "0.3"), results)
# results$fraction.pwb <- c("--", as.character(round(results$fraction.pwb[2], 0)))
# tab <- xtable::xtable(x=results, digits=1, align="rrrrrrrrrr", caption="Summary of results *1e5")

```

```{r save table, echo=FALSE, eval=FALSE}
ktab <- kbl(results,
            format="latex", 
            digits = 0, 
            align="rrrrrrrrrr",
            caption="Summary of results (multiplied by 100,000).",
            booktabs = T,
            col.names = c("$\\theta$", "Uncond.", "PWB", "", "$\\Delta$", "\\%Simon", "\\%SimonPWB", "Exc.", "Inc."),
            label="ma_add_trial",
            escape=FALSE)
ktab <- add_header_above(ktab, header = c(" "=1, "Bias"=2, "MCSE"=1, "Change in bias"=3, "MCSE"=2))
ktab <- add_header_above(ktab, header = c(" ", "Simon"=3, "Meta-analysis"=5))
ktab <- kable_styling(ktab, latex_options = c("hold_position", "scale_down"))
print(getwd())
#save_kable(ktab, file="simstudy12kbl.tex")
kable(ktab, format="html")
```


```{r create full bias data frame, message = FALSE}
simon.pw.bias.wtd <- simon.pw.bias*c(simon.frac, simon.frac2)
all.bias.uncond <- c(bias.nonAD, ma.bias.ex[1], simon.unwtd.bias[1], simon.frac/1e5,  
                     simon.frac*simon.unwtd.bias[1], prec.frac/1e5, prec.frac*simon.unwtd.bias[1], ma.bias.inc[1], 
                     bias.nonAD2, ma.bias.ex[2], simon.unwtd.bias[2], simon.frac2/1e5,
                     simon.frac2*simon.unwtd.bias[2], prec.frac2/1e5, prec.frac2*simon.unwtd.bias[2], ma.bias.inc[2])
all.pwb <- c(pwb.nonAD, NA, simon.pw.bias[1], NA, simon.frac*simon.pw.bias[1], NA, prec.frac*simon.pw.bias[1], NA,
             pwb.nonAD2, NA, simon.pw.bias[2], NA, simon.frac2*simon.pw.bias[2], NA, prec.frac2*simon.pw.bias[2], NA)
#wtd.pwb <- c(pwb.nonAD, ma.bias.ex[1], simon.frac*simon.pw.bias[1], ma.bias.inc[1],
#             pwb.nonAD2, ma.bias.ex[2], simon.frac2*simon.pw.bias[2], ma.bias.inc[2])
all.MCSE <- c(mcse.nonAD, ma.mcse.ex[1], simon.mcse[1], NA, NA, NA, NA, ma.mcse.inc[1],
             mcse.nonAD2, ma.mcse.ex[2], simon.mcse[2], NA, NA, NA, NA, ma.mcse.inc[2])
new.table <- data.frame(all.bias.uncond,
                        all.pwb,
                        all.MCSE)
new.table <- new.table*1e5
studies <- c("Study 1", "Study 2", "Study 3", "Study 4", "MA (exc.)", "Simon", "$W_i$", "$W_i$*Simon", "$p$", "$p$*Simon", "MA (inc.)",
             "Study 1", "Study 2", "Study 3", "Study 4", "MA (exc.)", "Simon", "$W_i$", "$W_i$*Simon", "$p$", "$p$*Simon", "MA (inc.)")
new.table <- cbind(studies, new.table)
names(new.table) <- c("Studies", "Unconditional", "PWB", "MCSE")
head(new.table, 8)
```

```{r, message = FALSE}
#options(knitr.kable.NA = '')
new.table.kbl <- kable(new.table)
new.table.kbl
#collapse_rows(kable_input=new.table.kbl, columns = 1, valign = "top")
new.table.kbl2 <- kable(new.table)
new.table.kbl2 <- kbl(new.table,
                      format="latex", 
                      digits = 2, 
                      align="rrrr",
                      caption="Detailed bias and MCSE results for individual studies and meta-analyses including and excluding adaptive (Simon) design (multiplied by 100,000). $W_i$: weight of Simon design as fraction of total meta-analysis. $p$: precision of Simon design as fraction of mean precision of non-adaptive designs",
                      booktabs = T,
                      col.names = c(" ", "Unconditional", "PWB", "MCSE"),
                      label="full.bias",
                      escape=FALSE)
new.table.kbl2 <- pack_rows(new.table.kbl2, "\\boldsymbol{$\\\\theta=0.5$}", 1, 11, escape = FALSE, bold = FALSE)
new.table.kbl2 <- pack_rows(new.table.kbl2, "\\boldsymbol{$\\\\theta=0.3$}", 12, 22, escape = FALSE, bold = FALSE)
new.table.kbl2 <- add_header_above(new.table.kbl2, c(" " = 1, "Bias" = 2, " " = 1))
new.table.kbl2 <- kable_styling(new.table.kbl2, latex_options = c("hold_position", "scale_down"))
new.table.kbl2
#save_kable(new.table.kbl2, file="simstudy12_table2.tex")
```

```{r smaller table}
bias.studies.uncond <- c(bias.nonAD, simon.unwtd.bias[1], bias.nonAD2, simon.unwtd.bias[2])
bias.studies.pwb <- c(pwb.nonAD, simon.pw.bias[1], pwb.nonAD2, simon.pw.bias[2])
studies.MCSE <- c(mcse.nonAD, simon.mcse[1], mcse.nonAD2, simon.mcse[2])

smaller.table <- data.frame(bias.studies.uncond, bias.studies.pwb, studies.MCSE)
smaller.table <- smaller.table*1e5

studies.coln <- rep(c("Study 1", "Study 2", "Study 3", "Study 4", "Simon"), times=2)
smaller.table <- cbind(studies.coln, smaller.table)

smaller.kbl <- kbl(smaller.table,
                      format="latex", 
                      digits = 1, 
                      align="rrr",
                      caption="Bias and MCSE for individual studies (multiplied by 100,000).",
                      booktabs = T,
                      col.names = c(" ", "Unconditional", "PWB", "MCSE"),
                      label="individual.bias",
                      escape=FALSE)
smaller.kbl <- pack_rows(smaller.kbl, "\\boldsymbol{$\\\\theta=0.5$}", 1, 5, escape = FALSE, bold = FALSE)
smaller.kbl <- pack_rows(smaller.kbl, "\\boldsymbol{$\\\\theta=0.3$}", 6, 10, escape = FALSE, bold = FALSE)
smaller.kbl <- add_header_above(smaller.kbl, c(" " = 1, "Bias" = 2, " " = 1))
smaller.kbl <- kable_styling(smaller.kbl, latex_options = c("hold_position", "scale_down"))
smaller.kbl
save_kable(smaller.kbl, file="bias_individual_studies.tex")
```

```{r bias in meta-analysis}
w.simon.uncond <- c(simon.frac*simon.unwtd.bias[1], simon.frac2*simon.unwtd.bias[2])
w.simon.pwb <- c(simon.frac*simon.pw.bias[1], simon.frac2*simon.pw.bias[2])
change <- ma.bias.inc-ma.bias.ex
abs.change <- abs(ma.bias.inc) - abs(ma.bias.ex)
fraction.OLD <- abs(change)/abs(simon.unwtd.bias)
fraction.pwb.OLD <- abs(change)/abs(simon.pw.bias)
fraction <- abs.change/abs(simon.unwtd.bias)
fraction.pwb <- abs.change/abs(simon.pw.bias)

bias.wide <- data.frame(1e5*ma.bias.ex, 
                        1e5*ma.mcse.ex,
                        1e5*ma.bias.inc,
                        1e5*ma.mcse.inc,
                        1e5*change,
                        1e5*abs.change,
                        fraction,
                        fraction.pwb, 
                        W=c(simon.frac, simon.frac2), 
                        1e5*w.simon.uncond, 
                        1e5*w.simon.pwb,
                        prec=c(prec.frac, prec.frac2))

#rownames(bias.wide) <- c("\\boldsymbol{$\\\\theta=0.5$}", "\\boldsymbol{$\\\\theta=0.3$}")
bias.wide <- cbind(c(0.5, 0.3), bias.wide)

summary.kbl <- kbl(bias.wide,
                      format="latex", 
                      digits = c(1, 1, 1, 1, 1, 1, 1, 2, 2, 2, 1, 1, 2), 
                      align="r",
                      caption="Detailed bias and MCSE results for meta-analysis (multiplied by 100,000) and $W_S$bias, where $W_S$ is the weight of the Simon design in the meta-analysis given $\\sum W_i=1$. ``Precision'' is the mean precision of Simon design as a factor of the overall mean precision of remaining studies.",
                      booktabs = T,
                      col.names = c("\\boldsymbol{$\\theta$}", "Bias", "MCSE", "Bias", "MCSE", "$\\Delta$", "$\\vert \\Delta_{inc} \\vert - \\vert \\Delta_{exc} \\vert$",  "Uncond.", "PWB", "$W_S$", "Uncond.", "PWB", "Precision"),
                      label="summary.bias",
                      escape=FALSE)
summary.kbl <- add_header_above(summary.kbl,
                                header=c(" "=1,
                                         "Excluding" = 2,
                                         "Including" = 2,
                                         " "=2,
                                         "$\\\\vert \\\\Delta_{inc} \\\\vert - \\\\vert \\\\Delta_{exc} \\\\vert$/$\\\\vert \\\\text{bias}$"=2,
                                         " "=1,
                                         "$W_S$bias"=2,
                                         " "=1),
                                escape = FALSE)
summary.kbl <- add_header_above(summary.kbl, c(" "=1, "Meta-analysis" = 6, "Simon" = 6))
summary.kbl <- kable_styling(summary.kbl, latex_options = c("hold_position", "scale_down"))
summary.kbl
save_kable(summary.kbl, file="bias_metas.tex")

```

