---
title: 'MA: Bias when adding trial'
author: "Martin Law"
date: '2022-06-06'
output: rmarkdown::html_vignette
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  dev='png',
  fig.path="figures/")
```

```{r setup}
library(pwb)
library(ggplot2)
```

We identify a trial design with substantial unadjusted bias and no precision weighted bias. We obtain bias for a meta-analysis before and after including this design. We expect no increase in (absolute) bias.


Find a Simon design with p~0~=0.5, p~1~=0.6, type-I error-rate 0.05, power
0.9:

```{r find_design}
all.designs <- clinfun::ph2simon(pu=0.5,
                                 pa=0.6,
                                 ep1=0.05,
                                 ep2=0.1,
                                 nmax=500)
# Choose optimal Simon design:
opt.index <- which.min(all.designs$out[, "EN(p0)"])
simon.des <- all.designs$out[opt.index, ]
```


Find bias for a true response probability theta=0.5:

```{r, find bias single }
# Find bias for this design, for a single true response probability theta:
  nsims <- 1e4
  theta <- 0.5
  set.seed(16)
  single.simon <- pwbSimon(theta=theta,
                                des=simon.des,
                                nsims=nsims)
  round(single.simon$ests[, 1:5], 5)
  round(single.simon$mc.error, 5)
```

```{r}
# Simulate all other MA studies:
ma.results <- simMeta(N=simon.des["n"], theta=0.5, n.studies=4, nsims=nsims)

# Summary estimate and bias (excluding Simon design):
  wtdmeans.ex <- rep(NA, nsims)
  for(i in 1:nsims){
    wtdmeans.ex[i] <- weighted.mean(x=ma.results$theta.hat[i, ], w=1/ma.results$se[i, ]^2)
  }
  bias.ex <- mean(wtdmeans.ex-theta)

```
Now add replications of the above Simon design to the meta-analysis:


```{r}
combined.results <- ma.results
combined.results$theta.hat <- cbind(single.simon$results$theta.hat, ma.results$theta.hat)
combined.results$se <- cbind(single.simon$results$se, ma.results$se)

# Summary estimate and bias (including Simon design):
  wtdmeans.inc <- rep(NA, nsims)
  for(i in 1:nsims){
    wtdmeans.inc[i] <- weighted.mean(x=combined.results$theta.hat[i, ], w=1/combined.results$se[i, ]^2)
  }
  bias.inc <- mean(wtdmeans.inc-theta)
  bias.ex
  bias.inc
  ### ^^^ looks like bias increases by almost an order of magnitude but still remains very small.
```




We then do the converse: identify a trial with 
