---
title: "Meta Analysis: Group Sequential"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MA_GS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  dev='png',
  fig.path="figures/")
```

```{r setup}
library(pwb)
```

Find a GS design with p0=0.5, p1=0.6, type-I error-rate 0.1, power 0.8:

```{r, gs}
gs.des <- curtailment::singlearmDesign(nmin=50,
                                   nmax=140,
                                   C=20,
                                   minstop=40,
                                   p0=0.5,
                                   p1=0.6,
                                   alpha=0.1,
                                   power=0.8,
                                   minthetaE=1,
                                   maxthetaF=0,
                                   max.combns=1e3)
# Obtain stopping bounds:
bounds <- curtailment::drawDiagram(gs.des)
```

Now simulate meta-analyses of a single GS design and some number of single-arm single-stage trials with the same maximum sample size _N_ as the GS design. We also choose the true response probability theta.

```{r}
set.seed(1453)
ma.gs.results <- maGS(theta=0.5,
                      des=gs.des,
                      bounds=bounds,
                      nsims=1e5,
                      n.studies=4)
ma.gs.results

```

What does the bias look like as theta is varied?

```{r, theta varied}
theta.vec.0.5 <- seq(0.2, 0.8, 0.1)
ma.gs.results0 <- vector("list", length(theta.vec.0.5))
for(i in 1:length(theta.vec.0.5)){
  ma.gs.results0[[i]] <- maGS(theta=theta.vec.0.5[i],
                              des=gs.des,
                              bounds=bounds,
                              nsims=1e4,
                              n.studies=4)
  }

ma.gs.results0 <- do.call(rbind, ma.gs.results0)
ma.gs.results0$p0p1 <- "p0=0.5, p1=0.6"

```


```{r "GS_MA_p05", fig.height=4, fig.width=4, echo=FALSE, warning=T}
library(ggplot2)
ma.gs.plot0 <- ggplot(data=ma.gs.results0, mapping=aes(x=theta, y=bias, col=type))+
  geom_line(alpha=0.4, size=1)+
  geom_point()+
  labs(x="True response probability theta",
       y="Bias (estimate - theta)",
       title="Bias in meta-analysis",
       subtitle="One GS design (p0=0.5, p1=0.6), four non-ADs")+
  geom_vline(aes(xintercept=0.5), col="grey", linetype="dashed")+
  geom_vline(aes(xintercept=0.6), col="grey", linetype="dashed")+
  geom_hline(aes(yintercept=0), col="grey", linetype="dashed")+
  ylim(max(abs(ma.gs.results0$bias)) * c(-1.1, 1.1) )+
  scale_x_continuous(breaks=theta.vec.0.5)+
  theme(legend.position="bottom")
ma.gs.plot0
```

### Moving away from theta=0.5

### p0=0.8

As with the single Simon design, let us examine what happens when the response rate is not close to 0.5. Here, we obtain a GS design for p0=0.8, p1=0.9. We reduce type-I error-rate to 0.05 and increase power to 90\% to increase the maximum sample size and number of interim analysis, both of which would be otherwise reduced by the high anticipated response rates.

```{r, gs high p}
gs.des2 <- curtailment::singlearmDesign(nmin=50,
                                   nmax=140,
                                   C=20,
                                   minstop=40,
                                   p0=0.8,
                                   p1=0.9,
                                   alpha=0.05,
                                   power=0.9,
                                   minthetaE=1,
                                   maxthetaF=0,
                                   max.combns=1e3)

# Obtain stopping bounds:
bounds2 <- curtailment::drawDiagram(gs.des2)
```

Again, simulate meta-analyses of this group sequential design and single-arm single-stage trials with the same maximum sample size _N_ as the group sequential design:

```{r, large thetea}
theta.vec <- seq(0.1, 0.9, 0.1)
ma.gs.results2 <- vector("list", length(theta.vec))
for(i in 1:length(theta.vec)){
  ma.gs.results2[[i]] <- maGS(theta=theta.vec[i],
                              des=gs.des2,
                              bounds=bounds2,
                              nsims=1e4,
                              n.studies=4)
  }

ma.gs.results2 <- do.call(rbind, ma.gs.results2)
ma.gs.results2$p0p1 <- "p0=0.8, p1=0.9"
```

```{r "gs_MA_p08", fig.height=4, fig.width=4, echo=FALSE, warning=T}
ma.gs.plot <- ggplot(data=ma.gs.results2, mapping=aes(x=theta, y=bias, col=type))+
  geom_line(alpha=0.4, size=1)+
  geom_point()+
  labs(x="True response probability theta",
       y="Bias (estimate - theta)",
       title="Bias in meta-analysis",
       subtitle="One GS design (p0=0.8, p1=0.9), four non-ADs")+
  #annotate("text", x=theta.vec2, y=-0.015, label=round(stop.early.count2,2))+
  geom_vline(aes(xintercept=0.8), col="grey", linetype="dashed")+
  geom_vline(aes(xintercept=0.9), col="grey", linetype="dashed")+
  geom_hline(aes(yintercept=0), col="grey", linetype="dashed")+
  ylim(max(abs(ma.gs.results2$bias)) * c(-1.1, 1.1) )+
  scale_x_continuous(breaks=theta.vec)+
  theme(legend.position="bottom")
ma.gs.plot
```

### p0=0.2

Obtain a GS design for p0=0.2, p1=0.3:

```{r, gs low p}
gs.des3 <-  curtailment::singlearmDesign(nmin=50,
                                   nmax=180,
                                   C=20,
                                   minstop=40,
                                   p0=0.2,
                                   p1=0.3,
                                   alpha=0.1,
                                   power=0.8,
                                   minthetaE=1,
                                   maxthetaF=0,
                                   max.combns=1e3)
# Obtain stopping bounds:
bounds3 <- curtailment::drawDiagram(gs.des3)
```

Again, simulate meta-analyses of this group sequential design and single-arm single-stage trials with the same maximum sample size _N_ as the group sequential design:

```{r, low theta}
theta.vec.low <- seq(0.1, 0.9, 0.1)
ma.gs.results3 <- vector("list", length(theta.vec.low))
for(i in 1:length(theta.vec.low)){
  ma.gs.results3[[i]] <-  maGS(theta=theta.vec.low[i],
                              des=gs.des3,
                              bounds=bounds3,
                              nsims=1e4,
                              n.studies=4)
  }

ma.gs.results3 <- do.call(rbind, ma.gs.results3)
ma.gs.results3$p0p1 <- "p0=0.2, p1=0.3"

# Biggest improvement in bias:
max(abs(ma.gs.results3$bias[ma.gs.results3$type=="All trials"])-abs(ma.gs.results3$bias[ma.gs.results3$type=="Exclude early stopped"]))
```

```{r "gs_MA_p02", fig.height=4, fig.width=4, echo=FALSE, warning=T}
ma.gs.plot.low <- ggplot(data=ma.gs.results3, mapping=aes(x=theta, y=bias, col=type))+
  geom_line(alpha=0.4, size=1)+
  geom_point()+
  labs(x="True response probability theta",
       y="Bias (estimate - theta)",
       title="Bias in meta-analysis",
       subtitle="One GS design (p0=0.2, p1=0.3), four non-ADs")+
  geom_vline(aes(xintercept=0.2), col="grey", linetype="dashed")+
  geom_vline(aes(xintercept=0.3), col="grey", linetype="dashed")+
  geom_hline(aes(yintercept=0), col="grey", linetype="dashed")+
  ylim(max(abs(ma.gs.results3$bias)) * c(-1.1, 1.1) )+
  scale_x_continuous(breaks=theta.vec.low)+
  theme(legend.position="bottom")
ma.gs.plot.low
```

```{r "gs_MA_combined", echo=FALSE, fig.height=4, fig.width=8, warning=TRUE}
ma.gs.results.comb <- rbind(ma.gs.results0, ma.gs.results2, ma.gs.results3)
#ma.simon.results0$p0p1 <- bquote(p[0]==0.5~","~p[1]==0.6)
ma.gs.plot.comb <- ggplot(data=ma.gs.results.comb, mapping=aes(x=theta, y=bias, col=type))+
  geom_line(alpha=0.4, size=1)+
  geom_point()+
  labs(x="True response probability theta",
       y="Bias (estimate - theta)",
       title="Bias in meta-analysis (group sequential design)")+
  geom_hline(aes(yintercept=0), col="grey", linetype="dashed")+
  theme(legend.position="bottom")+
  ylim(max(abs(ma.gs.results.comb$bias)) * c(-1.1, 1.1) )+
  scale_x_continuous(breaks=theta.vec.low)+
  facet_wrap(~p0p1, 
             scales = "free_x")#,
             #labeller=labeller(p0p1=letters[1:3]))
ma.gs.plot.comb
```

