---
title: 'Single studies: Fixed, Simon'
author: "Martin Law"
date: "`r Sys.Date()`"
output: html_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  dev='png',
  fig.path="figures/")
```



```{r setup}
library(pwb)
library(ggplot2)
library(ggrepel)
library(here)
source(here("R", "pwbSimonPlusNonAD.R"))
set.seed(19052025)
nsims <- 1e5
```


## Simulate Simon design with theta0=0.5
### ADEMP

- Aim: Compare conditional, unconditional, precision-weighted and UMVUE bias in a Simon design
- Data-generating mechanism: 
    * Binary outcome with a series of response rates (0.1, 0.9, 0.1)
    * Type-I error-rate 0.05 when response rate p~0~=0.5 and power 0.90 when response rate p~1~=0.6 
    * Corresponding design: N=233, n~1~=104. Trial failure if number of responses less than or equal to 54 at interim or less than or equal to 128 at end.
- Estimand: response rate
- Method of analysis: point estimate, SE
- Performance measures: Conditional, unconditional, precision-weighted and UMVUE bias
- Number of simulations: `r nsims`

Find a Simon design with p~0~=0.5, p~1~=0.6, type-I error-rate 0.05, power
0.9:

```{r find_design}
all.designs <- clinfun::ph2simon(pu=0.5,
                                 pa=0.6,
                                 ep1=0.05,
                                 ep2=0.1,
                                 nmax=300)
# Choose optimal Simon design:
opt.index <- which.min(all.designs$out[, "EN(p0)"])
simon.des <- all.designs$out[opt.index, ]
```



Find bias for a vector of true response rates:
```{r, find bias vector, cache=F}
##### Obtain bias for range of true response probabilities ####
theta.vec <- seq(0.1, 0.9, 0.05) # Vector of true response probabilities
summary.data <- raw.data <- mc.error <- vector("list", length(theta.vec))

for(i in 1:length(theta.vec)){
  one.run <- pwbSimonPlusNonAD(theta=theta.vec[i],
                           des=simon.des,
                           nsims=nsims)
  raw.data[[i]] <- one.run$results
  summary.data[[i]] <- one.run$ests
  mc.error[[i]] <- one.run$mc.error
}
single <- do.call(rbind, summary.data)
```


```{r}
# Plot or print proportion of early-stopped trials, excluding extremes:
single$early.stop.subset <- ifelse(single$early.stop>0.01 & single$early.stop<0.99 & single$Design=="Simon" & single$Bias=="Unconditional", yes=single$early.stop, no=NA)

single$early.stop.subset <- ifelse(!is.na(single$early.stop.subset), yes=sprintf("%.2f", single$early.stop.subset), no="")
```

Plot results
```{r "single_studies", echo=FALSE, fig.height=6, fig.width=8, warning=TRUE}
single.plot <- ggplot(data=single, mapping=aes(x=theta, y=bias, col=Bias, linetype=Design))+  geom_line(alpha=0.4, linewidth=1)+
  geom_point()+
  labs(x="True response probability theta",
       y="Bias",
        title=bquote("Bias in single studies ("*theta[0]==0.5*", "~theta[1]==0.6*")"),
       subtitle="Includes proportion of trials stopped early (selected)")+
  geom_text_repel(mapping=aes(label=early.stop.subset), colour="black", size=5, seed=1)+
  geom_vline(aes(xintercept=0.5), col="grey40", linetype="dashed")+
#  geom_vline(aes(xintercept=0.6), col="grey40", linetype="dashed")+
  geom_hline(aes(yintercept=0), col="grey40", linetype="dashed")+
 # ylim(max(abs(simon.sub.df$bias)) * c(-1.2, 1.2) )+
  scale_x_continuous(breaks=seq(0.1,0.9,0.1))+
 # scale_color_manual(values=simon.colours)+
  theme(legend.position="bottom",
        legend.box="vertical")+ 
  theme(text=element_text(size=20))
single.plot

ggsave(filename="single_studies.eps",
       plot=single.plot,
       device=cairo_ps,
       path=here("vignettes", "figures"))
```


## Simulate trials again, using values of p~0~ and p~1~ that are not close to 0.5

```{r "simon_high_p"}
all.designs2 <- clinfun::ph2simon(pu=0.8,
                                 pa=0.9,
                                 ep1=0.05,
                                 ep2=0.1,
                                 nmax=150)
# Choose optimal Simon design:
opt.index <- which.min(all.designs2$out[, "EN(p0)"])
simon.des2 <- all.designs2$out[opt.index, ]
```



Find bias for a vector of true response rates:
```{r "find bias vector 2", cache=F}
theta.vec <- seq(0.1, 0.9, 0.05) # Vector of true response probabilities
summary.data <- raw.data <- mc.error <- vector("list", length(theta.vec))

for(i in 1:length(theta.vec)){
  one.run <- pwbSimonPlusNonAD(theta=theta.vec[i],
                           des=simon.des2,
                           nsims=nsims)
  raw.data[[i]] <- one.run$results
  summary.data[[i]] <- one.run$ests
  mc.error[[i]] <- one.run$mc.error
}
single <- do.call(rbind, summary.data)
```



```{r}
# Plot or print proportion of early-stopped trials, excluding extremes:
single$early.stop.subset <- ifelse(single$early.stop>0.01 & single$early.stop<0.99 & single$Design=="Simon" & single$Bias=="Unconditional", yes=single$early.stop, no=NA)

single$early.stop.subset <- ifelse(!is.na(single$early.stop.subset), yes=sprintf("%.2f", single$early.stop.subset), no="")
```


Plot results
```{r "single_studies2", echo=FALSE, fig.height=6, fig.width=8, warning=TRUE}
single.plot2 <- ggplot(data=single, mapping=aes(x=theta, y=bias, col=Bias, linetype=Design))+  geom_line(alpha=0.4, linewidth=1)+
  geom_point()+
  labs(x="True response probability theta",
       y="Bias",
        title=bquote("Bias in single studies ("*theta[0]==0.8*", "~theta[1]==0.9*")"),
       subtitle="Includes proportion of trials stopped early (selected)")+
  geom_text_repel(mapping=aes(label=early.stop.subset), colour="black", size=5, seed=1)+
  geom_vline(aes(xintercept=0.8), col="grey40", linetype="dashed")+
#  geom_vline(aes(xintercept=0.6), col="grey40", linetype="dashed")+
  geom_hline(aes(yintercept=0), col="grey40", linetype="dashed")+
 # ylim(max(abs(simon.sub.df$bias)) * c(-1.2, 1.2) )+
  scale_x_continuous(breaks=seq(0.1,0.9,0.1))+
 # scale_color_manual(values=simon.colours)+
  theme(legend.position="bottom",
        legend.box="vertical")+ 
  theme(text=element_text(size=20))
single.plot2

ggsave(filename="single_studies2.eps",
       plot=single.plot2,
       device=cairo_ps,
       path=here("vignettes", "figures"))
```
